\section{\noindent \label{sec:Introduction}Introduction}

Type II diabetes mellitus is increasing worldwide, but especially
in \ac{LMICs} such as Mexico \citep{InternationalDiabetesFederation2013}.
This increase in prevalence is likely driven by changes in diet and
physical activity \citep{Barquera2008b,Basu2013} and a distinct genetic
predisposition of Mexicans to develop \ac{T2D} \citep{Williams2013}.
Apart from the health consequences of \ac{T2D}, it also has economic
consequences causing a considerable amount healthcare expenditures
and possibly affecting peoples employment chances and productivity. 

In the literature many studies have found a negative effect of diabetes
on employment chances and wages, though the evidence from \ac{LMICs}
remains limited \citep{Seuring2015a}. Diabetes could affect labour
market outcomes via several pathways: People with diabetes could drop
out of employment due to the health consequences of diabetes making
them less productive than their healthy counterparts. Further, employers
could discriminate against people with diabetes deeming them to be
more risky to employ than people without diabetes due to their health
problems and greater need for medical treatment. Finally, people aware
of their diabetes might also be less inclined to continue working
if this interferes with their effective management of the disease.
The last two points possibly also could result in different effects
for people with diagnosed versus people with undiagnosed diabetes
despite both of the having a similar state of health. Also, diabetes
is not a homogeneous disease in the sense that it can have different
effects on people's health depending on the success of the long term
disease management. On the one hand people who are able to ''reverse''
their diabetes, meaning they managed to get back to healthy blood
glucose levels as a result of lifestyle chances that successfully
reestablish normal insulin sensitivity, are unlikely to suffer from
any diabetes related health problems \citep{Lim2011,Gregg2012}. On
the other hand, if diabetes is either completely untreated or unsuccessfully
treated by medicine or lifestyle changes causing blood glucose levels
to stay elevated and insulin resistance high over a longer period
of time, people tend to develop various debilitating health conditions
as time progresses \citep{Reynoso-Noveron2011}. The goal of this
study is to investigate these points using data from the \acf{MxFLS}
to provide evidence for the relationship of self-reported diabetes,
the time since diagnosis, undiagnosed diabetes as well as the severity
of diabetes with labour market outcomes.

Various studies have investigated the labour market impact of diabetes.
Most of them have focused on high-income countries and little is know
on the relationship of diabetes with labour market outcomes in low-
and middle-income countries with only two studies providing evidence
so far. \citet{Liu2014} investigate the effect of a recent diabetes
diagnosis on labour income in China using a natural experiment for
identification and find a significant reduction in income for those
with a recent diagnosis. For Mexico, in an earlier paper we investigate
the effect of self-reported diabetes on employment chances using cross-sectional
data from the 2005 wave of the \ac{MxFLS}. We find a significant
(p<0.01) reduction in employment chances for males of a about 10 percentage
points and for females of about 4.5 percentage points, albeit at a
much lower statistical significance \citep{Seuring2015}. 

Regarding the duration of diabetes, using pooled panel data from an
US population a logit and Heckman type model, \citet{Minor2013} finds
some evidence for a non-linear relationship of \ac{T2D} duration
and employment two to five years after diagnosis for men and eleven
to fifteen years after diagnosis for women. He finds, however, very
limited evidence for a linear relationship with employment and for
any relationship with wages. One study by \citet{BrownIII2011} looked
at how diabetes management inferred from measured \ac{HbA1c} levels
affected employment chances and wages using cross-sectional data in
a mainly Mexican-American population in the US. Also using a probit
and a Heckman type model they found a strong adverse association of
a linear increase in \ac{HbA1c} levels with employment chances and
with wages for men.

Other studies have been limited in their scope to self-reported diabetes
and only estimated the average effect of diabetes on employment chances
or wages not investigating the effects of diabetes duration. Further,
these studies have used only cross-sectional data. Because there is
the possibility that the estimates of the effects of self-reported
diabetes might suffer from omitted variable bias, many studies have
used an \ac{IV} strategy to achieve identification of the true relationship
of diabetes and labour market outcomes. In the case of diabetes and
labour market outcomes, personal characteristics such as time preference
and native ability, genetic trades, poor early life health including
in the mother' womb, or low income in early life have been shown to
adversely affect health and more specifically the propensity to develop
type 2 diabetes \citep{VanEwijk2011a,Sotomayor2013,Li2010b}, and
could effect employment chances, wages or working hours indirectly
through their adverse effects on educational attainment \citep{Ayyagari2011b}
as well as directly through their effects on today's productivity
\citep{Currie2013}. The instrumental variable strategy has thereby
remained the same across studies with all of them using the family
history of diabetes as their instrument, relying on heritable genetic
trades of diabetes that increase the chances of developing diabetes
if one of the parents has or has had the disease \citep{Brown2005,Latif2009,Minor2010a,Lin2011b,Seuring2015}.
However, some drawbacks to the use of instruments exist: first it
is not readily testable if the instrument is uncorrelated with the
outcome of interest other than through its causal relationship with
the endogeneous explanatory variable, and second the use of instruments
results in biased estimates in finite samples and generally produce
large standard errors particularly if the instrument is only weakly
correlate with the endogenous variable \citep{Bound1995}. 

Because other plausible instruments for diabetes are difficult to
find, another method to account for unobserved omitted variables is
the use of individual level \ac{FE}. To do this repeated observations
of the same individual over time are needed which are only available
in panel data. The use of this method allows for differencing out
the fixed time invariant unobservables that could potentially bias
our estimates. This includes all of the above mentioned possible confounders
related to genes and early life environments.

Thanks to the three waves provided by the \ac{MxFLS} we are able
to use this method for the first time in this context, to investigate
how self-reported diabetes and diabetes duration are associated with
labour market outcomes. Apart from this contribution, our paper makes
several other important contributions to the literature. First, we
extend the analysis to wages and working hours, a topic not yet investigated
in the context of Mexico. Second, we investigate the heterogeneity
in the relationship of self-reported diabetes and its relationship
with wages and working hours across different employment types, i.e.
non-agricultural employment, agricultural employment and self-employment.
This is important due to the large agricultural and self-employed
sectors, where health problems might have much different effects than
in more secure or less strenuous non-agricultural jobs. Third, because
we have information from three different points in time we are able
to investigate the extend of inconsistent self-reported diabetes in
the data and to correct for it to increase the validity of our results.
Finally, we provide a first exploration of the labour market effects
of undiagnosed diabetes, which is important as a large part of the
population with diabetes is unaware of the disease preventing early
treatment to prevent the adverse health effects of diabetes.

The remainder of the paper is structured as follows. Section \ref{sec:METHODOLOGY}
presents the used data, its inconsistencies, discusses the problem
of measurement error in our data and lays out the empirical strategy.
Section \ref{sec:RESULTS} provides the results of the panel data
analysis and the cross-sectional biomarker analysis. Section \ref{sec:Conclusion}
discusses the results and concludes.

\section{Methodology}
\subsection{The data}
The dataset used for the empirical analysis is the \acf{MxFLS}.
It is a nationally representative longitudinal household survey with
currently three waves conducted in 2002, 2005--2006 and 2009--2011.
Interviews were conducted with all household members aged 15 and above
and information on a wide range of social, demographic, economic characteristics
and health behaviours of the individuals and their families was collected
\citep{Rubalcava2013}. Apart from self-reported diabetes information
throughout the survey, the third wave also provides biomarker data
for everybody age 45 and above and for a random subsample of individuals
below 45 years. The third wave also provides information on the self-reported
year of diagnosis, information that was missing from the previous
two waves. The sample we use is restricted to the working age population
between age 15 and 64. For our main analysis using self-reported diabetes
we make use of all three waves in order to take advantage of the large
amount of observations and the panel data structure of the data.
\subsection{Data inconsistencies}

An apparent issue in the use of self-reported data is reporting error.
However, one of the advantages of panel data is that it provides repeated
measures for many of the individuals, allowing to uncover inconsistencies
and the extend thereof for those individuals with at least two observations.
While we could not find any literature investigating this issue for
self-reported diabetes, a study by \citet{Zajacova2010} on the consistency
of a self-reported cancer diagnosis over time in a US population found
that 30 percent who had reported a cancer diagnosis at an earlier
point, reported at a later point that they never had received a cancer
diagnosis. They also found that a more recent diagnosis was reported
with greater consistency possibly due to increasing recall problems
or lower salience as time since diagnosis advances. 

We find similar inconsistencies in the diabetes self-reports over
the three waves of the \ac{MxFLS} data, with between 10 to 20 percent
of those reporting diabetes in one wave not reporting diabetes in
the one of the ensuing waves. While we could not find a study assessing
the validity of self-reported diabetes in Mexico, a study from China
has shown that specificity -meaning the correct self-report of a diabetes
diagnosis- was very high (>98 percent for China), while sensitivity
-that is the awareness of the disease- was low (40 percent for China)
\citep{Yuan2015a}. This indicates that people who report a diagnosis
of diabetes are likely to have it while many of those reporting no
diagnosis have diabetes as well but are not diagnosed. 

Thanks to the biomarker data provided in the third wave of the \ac{MxFLS},
we are also able to assess the validity of self-reported diabetes
by using \ac{HbA1c} levels as well as the report of diabetes related
medicine use as a confirmation of self-reports. The \ac{WHO} recommends
a cut-off value of an \ac{HbA1c} $\geq6.5$percent, to diagnose
a person with diabetes \citep{WorldHealthOrganization2011}. Of the
subsample selected for biomarker measurements and answering the diabetes
question (n=6895), 705 reported a diabetes diagnosis and of those
also 632 (90\%) had an \ac{HbA1c} $\geq6.5$percent or reported
taking diabetes medication. Further, the biomarker data shows that
of the 1803 having an \ac{HbA1c} $\geq6.5$percent, 1265 did
not self-report a diabetes diagnosis in the third wave, indicating
that 70 percent of the people with diabetes are undiagnosed in the
subsample. These results support those from China indicating that
diabetes self-report has a high specificity but a low sensitivity.

Due to the high specificity in the reporting of a diabetes diagnosis,
we assume for people with information from only two waves, that if
they reported a diabetes diagnosis in a prior wave they also had diabetes
in the ensuing wave even if they did not report a diabetes diagnosis.
For people where we had information on all three waves, we used that
additional information to make a decision on how to deal with inconsistencies
using the following rules (Table \ref{tab:Inconsistencies}):
\providecommand{\tabularnewline}{\\}

\begin{table}[h!]
\begin{center}
%\resizebox{\textwidth}{!}{%
\begin{tabular}{llc}
Inconsistency & Assumption & Number of observations replaced\tabularnewline
\hline 
Diabetes self report in 2002, 2005 but not in 2009 & Has diabetes in 2009 as well & 34\tabularnewline
Diabetes self report in 2002, 2009 but not in 2005 & Has diabetes in 2005 as well & 94\tabularnewline
Diabetes self report only in 2002, but not in 2005 and 2009 & Has no diabetes in 2002 as well & 86\tabularnewline
Diabetes self report only in 2005, but not in 2002 and 2009 & Has no diabetes in 2005 as well & 71\tabularnewline
Diabetes self report in 2002, but not in 2005. Not in survey in 2009 & Has diabetes in 2005 as well & 43\tabularnewline
Diabetes self report in 2005, but not in 2009. Not in survey in 2002 & Has diabetes in 2009 as well & 32\tabularnewline
\hline 
\end{tabular}
%}
\end{center}
\caption{\label{tab:Inconsistencies}\textbf{Inconsistencies in diabetes self-report in MxFLS}}
\end{table}
  

This approach should add more consistency to the self-reported diabetes
information by using all available information. We tested if the results
of the \ac{HbA1c} tests for the subpopulation in 2009 with biomarker
information and inconsistencies in their diabetes reporting (n=96,
48 with two and 48 with one self-report of diabetes) would support
this decision by comparing the mean \ac{HbA1c} values for those who
had two self-reports of a diagnosis of diabetes in the full three
waves with those with only one self-report. Using a t-test, we find
a significantly (p<0.001) higher mean \ac{HbA1c} of 9.6 percent for
those with two self-reports compared to 7.0 percent for those with
only one self-report of diabetes. Further, of those with one self-report,
only 30 percent had an \ac{HbA1c}$\geq6.5$percent compared
to 87 percent of those with two self-reports. As a result the way
we have dealt with the inconsistencies in the data should minimize
misclassification of people into diabetes or no-diabetes and should
reduce some of the measurement error in the diabetes data. 

Unfortunately we cannot use a similar method for dealing with inconsistencies
in the self-reported year of diabetes diagnosis, which is why we have
to rely on the given information. Taking this into account the results
from models using duration of diabetes instead of self-reported diabetes
must be interpreted very carefully.


\subsection{Some more on measurement error for self-reported diabetes}

Self-reported data can suffer from severe non-classical measurement
error due to systematic misreporting (mostly underreporting in the
case of \ac{BMI} or obesity) which has been shown to cause estimates
of economic outcomes to be biased and overstated \citep{Cawley2015,ONeill2013,Perks2015}.
For diabetes, specifying measurement error is somewhat more difficult
as it needs to be clear what kind of measurement is used to detect
diabetes in survey data.

When talking about diabetes in the context of most survey data it
normally means referring to a self-reported diabetes diagnosis. In
the context of the \ac{MxFLS}, survey participants where asked ''Have
you ever been diagnosed with diabetes?''. The response to this question
likely suffers from some measurement error, depending on how well
the respondent understands the question and possibly how long ago
a diagnosis might has been made. However, as discussed earlier, specificity
for self-reported diabetes is quite high in general and also in the
\ac{MxFLS}. The main issue potentially causing measurement error
is overreporting, as about ten percent report a diabetes diagnosis
without having an \ac{HbA1c} above the diabetes threshold and at
the same time do not take any diabetes medication. It is difficult
to quantify the direction of this bias. It could attenuate the adverse
effects of diabetes if the whole group would not have diabetes and
consequently also experience no adverse health effects due to diabetes.
However, if some of them falsely reported a diabetes diagnosis to
justify something else in their life such as their current unemployment,
this could potentially lead to an overestimation of the true effect
of diabetes on employment chances in this case. Lastly, there is the
possibility that many of those having an \ac{HbA1c} $<6.5$ percent
have received a diagnosis and are successfully managing their diabetes
by lifestyle changes alone without the need for medication. This would
cause no measurement error at all. 

Overall, while measurement error can be a problem for diabetes research,
one has to be careful in defining what is being examined: self-reported
diabetes or diabetes as in those with biomarkers above a certain threshold.
For the former, measurement error should be rather small as most people
accurately report a diagnosis. For the latter, measurement error depends
on the accuracy of the used biomarker test as to how accurate they
are in measuring the relevant information.

Nonetheless, if we are interested in the economic effects of diabetes
and use self-reported diabetes as our diabetes indicator, we must
be aware that we are likely measuring something else then the pure
medical effects of diabetes. A diabetes diagnosis possibly also affects
a person's psychology and health behaviour which could have its own
effects on economic outcomes.\footnote{For example \citet{Liu2014} found that receiving a diabetes diagnosis
considerably reduced labour income in Chinese employees shortly after
their diagnosis. Similarly, others have shown that a hypertension
diagnosis can considerably affect health behaviours \citep{Zhao2013a}.
Similar effects have also been found for the US that people receiving
a diabetes diagnosis changed their health behaviours favourably albeit
only over the short term \citep{Slade2012}. Such changes in health
behaviours after a diabetes diagnosis could also translate into changes
in employment chances or productivity.} Accordingly, the effects we measure with self-reported diabetes are
likely different from those that we would measure based on a purely
medical definition of diabetes assessed via blood tests.


\subsection{Estimation strategy}

The fixed effect model then deals with the problem of unobserved time-invariant
confounders by taking away the group mean for each individual from
both sides of the estimated equation:

\begin{equation}
y_{it}-\overline{y_{i}}=\beta(X_{it}-\overline{X_{i}})+(\alpha_{i}-\overline{\alpha_{i}})+(u_{it}-\overline{u_{i}}),\label{eq:fixed}
\end{equation}


where $y_{it}$ is the dependent variable observed for individual
$i$ at time $t$, $X_{it}$ represents the observed time-variant
explanatory variables, $\alpha_{i}$ is the unobserved time-invariant
individual effect and $u_{it}$ is the error term. As equation \ref{eq:fixed}
shows, the unobserved time-invariant effects $\alpha_{i}$ are differenced
out as they do not vary over time. Only the variation in time variant
variables is left to be estimated so that changes in the dependent
variable are explained by changes in the dependent variables. This
will be the main way we investigate the relationship of self-reported
diabetes with labour market outcomes. While using fixed-effects estimation
does not identify a causal relationship as other forms of heterogeneity
could affect our estimates, i.e. time-variant unobserved heterogeneity
or simultaneity, it gives us a better idea of the relationship than
simple cross-sectional data and does allow to control for unobserved
personal characteristics that could bias the estimates without the
drawbacks of an \ac{IV} strategy. 




\subsection{The specification}

We estimate the relationship of self-reported diabetes and three labour
market outcomes, i.e. employment chances, log hourly wages  and normal
weekly working hours using two-way fixed effects for all observations
as well as stratified by gender. We estimate a two-part model where
the model for employment chances takes the following form:

\noindent 
\begin{equation}
Employed_{it}=\beta_{0}+\beta_{1}Diabetes_{it}+\beta_{2}X_{it}+c_{i}+\gamma_{t}+u_{it}.\label{eq:employed}
\end{equation}


where $Y_{it}$ is a binary variable taking a value of $1$ if respondent
$i$ reports being employed at time $t$ and $0$ otherwise, $Diabetes_{it}$
is a binary variable taking a value of $1$ if the respondent reports
having ever received a diagnosis of diabetes, $X_{it}$is a vector
of control variables, $c_{i}$ represents an individual fixed effect,
$\gamma_{t}$ represents a year fixed effect, and $u_{it}$ is the
error term.

For the relationship of self-reported diabetes with log hourly wages
 and weekly working hours models are estimated conditional on being
employed:

\noindent 
\begin{equation}
Y_{it}=\beta_{0}+\beta_{1}Diabetes_{it}+\beta_{2}X_{it}+c_{i}+\gamma_{t}+u_{it}.\label{eq:income_working-hours}
\end{equation}


where $Y_{it}$ is the log hourly wage of respondent $i$ at time
$t$ or in the case of working hours the usual weekly working hours
over the last year. The control variables in both specifications include
age and age squared to capture any non-linearities between age and
labour market outcomes, dummy variables for the effects of any changes
in the living environment of living in a small, medium or large city
with rural and regional dummies for five different Mexican regions\footnote{The region variables have been constructed after recommendations on
the MxFLS-Homepage. South-southeastern Mexico: Oaxaca, Veracruz, Yucatan;
Central Mexico: Federal District of Mexico, State of Mexico, Morelos,
Puebla; Central northeast Mexico: Coahuila, Durango, Nuevo Leon; Central
western Mexico: Guanajuato, Jalisco, Michoacan; Northwest Mexico:
Baja California Sur, Sinaloa, Sonora.} with living in the northeast being the base. We also include a martial
status dummy to control for the impact of marriage on employment chance.
A variable capturing the number of children residing in the household
below the age of 18 is used to control for the impact of children
on labour market outcomes and the effect of childbearing and related
gestational diabetes on the probabilities of developing type 2 diabetes
\citep{Bellamy2009}. To account for the effect that changes in household
wealth might have on diabetes and employment chances, we use the well
established method of principal component analysis of multiple indicators
of household assets and housing conditions to create an indicator
for household wealth \citep{Filmer2001}. The components used for
the construction of the wealth index are detailed in \citet{Seuring2015}.
Year fixed effects are included to capture any changes in the macroeconomic
environment in Mexico over the observed period that could affect employment
chances, wages or working hours and simultaneously the chances to
develop diabetes.

Additionally, we reestimate the above models using a self-reported
measure of the years since diagnosis. First, using a linear specification
indicating the association of our outcomes with each additional year
since diagnosis. Second, using linear splines of time since diagnosis
to capture non-linearities in the relationship of diabetes duration
and labour market outcomes. Because the year of diagnosis was only
reported in the third wave, we could only calculate the duration of
diabetes (or time since diagnosis) for the earlier waves for those
that had also responded to the third wave. To arrive at the time passed
since diagnosis, the year of diagnosis was subtracted from the year
of the interview. Those that reported a diagnosis in the year of the
interview were counted as one year since diagnosis. Accordingly, if
the respondent reported to having been diagnosed in the year before
the interview he was counted as two years since diagnosis. This was
done to have zero years representing people without diabetes and to
use them as the reference category.

Finally, we make use of the biomarker data in wave three, which allows
the exploration of the relationship of undiagnosed diabetes as well
as diabetes severity with labour market outcomes, as measured by \ac{HbA1c}
values. Because this data is only available in the last wave and it
is not possible to use the information in a meaningful way for the
earlier waves, we have to rely on cross-sectional methods. This has
obviously drawbacks in that we are not able to account for unobserved
heterogeneity using fixed effects. Further, as mentioned before, biomarkers
were only taken from a subsample of about 1/3 of our original sample
for the third wave, which leaves us with information on 6994 survey
participants. Therefore the analyses cannot be directly compared to
the other results in this paper. Nonetheless, in allows for a first
exploration of the relationships of undiagnosed diabetes as well as
disease severity with labour market outcomes. This is important as
the undiagnosed represent a large part of the diabetes population
in Mexico which has not been explored so far. It further allows us
to explore some of the heterogeneity that exists within the diabetes
population in terms of how well people's diabetes is managed.


\subsection{Sample characteristics}

Before moving on to the empirical results we will take a look at the
sample we use for analysis stratified by men and women.

As visible in Figure \ref{fig:Self-reported-diabetes-prevalenc},
unweighted self-reported diabetes prevalence in the \ac{MxFLS} has
increased from about 6 percent in 2002 to 7.1 percent in 2009 for
females and from about 4.3 to 5.7 percent for males. This is still
well below the estimated prevalence of diabetes by other institutions
such as the \ac{IDF}, whose most recent estimates for 2014 indicate
a prevalence of about 12 percent (equalling circa 9 million Mexicans)
for those aged between 20 and 79 \citep{InternationalDiabetesFederation2013}.
They further estimate that about 25 percent of the diabetes population
are not aware of their diabetes. This difference in self-reported
and undiagnosed diabetes should, together with the different age groups
used, explain the differences between the self-reported diabetes prevalence
in the \ac{MxFLS} and the actual diabetes prevalence in Mexico.  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  