\subsubsection*{Diagnosed and undiagnosed diabetes}

Table \ref{tab:Self-reported-diabetes-and-1}
indicates no association between undiagnosed
diabetes and any labour outcome. For those with diagnosed
diabetes the coefficients for the association with employment chances
are similar in size and significance for the pooled sample and women,
compared to those of the fixed effects model shown in table \ref{tab:Self-reported-diabetes-and} and to the cross-sectional model only including self-reported diabetes shown in table \ref{tab:Self-reported_diabetes_biomarkersample}.
For men, the coefficient is again smaller and not statistically significant
at conventional levels (p=0.103). We further find no indication of strong associations
between any form of diabetes and wages or working hours. 
Overall, the finding that undiagnosed diabetes appears to not be associated with any adverse labour outcomes is rather puzzeling as it suggests that the two populations with diabetes, those with diagnosed and those with undiagnosed diabetes are different in terms that are not yet captured by the used control variables. Possible differences that could explain these different outcomes could be the psychological effects of a diabetes diagnosis leading to a reduction in labour supply over time. Further, people with a diabetes diagnosis could be generally less healthy and have more diabetes related complications or other diseases that affect their employment decision. However, accounting for BMI, self-reported hypertension and self-reported heart disease as well as a measure of subjective health does only marginally affect the coefficients for self-reported diabetes and undiagnosed diabetes, reducing the former by about one percentage point (results not shown here). In a further exploration we will need to see if accounting for psychological well being and other potential confounders will effect the estimates and explain some of the differences in the observed effects.

\begin{table}[h]
\begin{center}
%\resizebox{\textwidth}{!}{%
{ \def\sym#1{\ifmmode^{#1}\else\(^{#1}\)\fi} \begin{tabular}{l*{9}{D{.}{.}{-1}l}} \toprule&\multicolumn{3}{c}{Employment}&\multicolumn{3}{c}{Log hourly wages}&\multicolumn{3}{c}{Weekly working hours}\\\cmidrule(lr){2-4}\cmidrule(lr){5-7}\cmidrule(lr){8-10}&\multicolumn{1}{c}{(1)}&\multicolumn{1}{c}{(2)}&\multicolumn{1}{c}{(3)}&\multicolumn{1}{c}{(4)}&\multicolumn{1}{c}{(5)}&\multicolumn{1}{c}{(6)}&\multicolumn{1}{c}{(7)}&\multicolumn{1}{c}{(8)}&\multicolumn{1}{c}{(9)}\\&\multicolumn{1}{c}{Pooled}&\multicolumn{1}{c}{Males}&\multicolumn{1}{c}{Females}&\multicolumn{1}{c}{Pooled}&\multicolumn{1}{c}{Males}&\multicolumn{1}{c}{Females}&\multicolumn{1}{c}{Pooled}&\multicolumn{1}{c}{Males}&\multicolumn{1}{c}{Females}\\
\midrule Diagnosed diabetes&    -.061\sym{***}&    -.036         &    -.052\sym{**} &     .005         &     .020         &    -.021         &    -.829         &    -.234         &   -1.938         \\
                &   (.017)         &   (.026)         &   (.023)         &   (.054)         &   (.063)         &   (.103)         &  (1.140)         &  (1.329)         &  (2.168)         \\
\addlinespace
Undiagnosed diabetes&    -.005         &     .013         &    -.016         &    -.032         &    -.039         &    -.045         &    1.236         &     .792         &    2.399         \\
                &   (.013)         &   (.016)         &   (.020)         &   (.038)         &   (.045)         &   (.068)         &   (.793)         &   (.880)         &  (1.571)         \\
\midrule
R2              &     .325         &     .074         &     .138         &     .256         &     .247         &     .303         &     .073         &     .047         &     .050         \\
N               &     6476         &     2786         &     3690         &     2694         &     1802         &      892         &     3455         &     2301         &     1154          \\ \bottomrule \multicolumn{10}{l}{\footnotesize Robust standard errors in parentheses}\\ \multicolumn{10}{l}{\footnotesize Other control variables: age, region, urban, education, indigenous, marital status, children, wealth, gender.}\\ \multicolumn{10}{l}{\footnotesize \sym{*} \(p<0.10\), \sym{**} \(p<0.05\), \sym{***} \(p<0.01\)}\\ 
\end{tabular}%
}
\end{center}
\caption{\label{tab:Self-reported-diabetes-and-1}\textbf{Self-reported diabetes and undiagnosed diabetes and labour market outcomes}}
\end{table}
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  